\subsection{Fault Tolerance}
\label{sec:fault-tolerance}

Migrations in Shadowfax can be easily made fault tolerant.
%
% In addition to marking their part of the migration completed at the
% metadata store
During their
respective \texttt{Complete} phases in the protocol,
the source and target would first have to take a checkpoint before
setting their flags at the metadata store.
%
This would make the migration durable;
% , and the temporary dependency
% between the source and target at the metadata store is garbage collected.
%
if either machine crashes hereafter, it can be independently recovered
from a checkpoint containing the effects of the migration.

If either server crashes during, recovery must involve both, which is why the
metadata store tracks the dependency between them.
%
This is because of early ownership transfer; during migration, the
target services operations on the migrating ranges, but many records belonging
to it may still be on the source.
%
When recovering a server, if Shadowfax finds a migration dependency
involving the server without both completion flags set, it cancels the
migration by setting a cancellation flag in the metadata server.
%
Then, it transfers ownership of hash ranges back to the source
(incrementing the source and target's view), restores both machines using
their pre-migration checkpoints, and recovers requests on hash
ranges that were issued during migration at the source.

This cancellation procedure ensures migration is deadlock-free.
%
If either server fails to make progress through the protocol
in a timely manner, the migration can always be cancelled by any party, and
both servers can be rolled back.
%
No server can stall migration completion indefinitely.

%
%As with ownership transfer, it creates a global cut across client
%sessions, indicating which operations clients should retransmit
%to servers after a crash.
%
%This client-assisted recovery eliminates the need for write-ahead-logging at
%servers, which would introduce a serial bottleneck.
%
