\section{Discussion}
\label{sec:discussion}

Shadowfax's techniques are not restricted to KVSs and can be applied to
other systems as well.
%
Its partitioned sessions can be used by \emph{stateful} cloud services
to preserve throughput over the network.
%
In fact, our implementation of sessions is templated on the service; we
used \faster for the purpose of this system, but one could also use
parameter servers, graph stores, model serving systems etc.

Likewise, asynchronous global cuts can be used to scale-out these
services while preserving throughput.
%
Since these cuts help propagate changes in ownership
across cores and machines, they can also be used for other operations
that involve changes in ownership like
failure detection and crash recovery.

In addition to scale-out, Shadowfax's migration protocol can also be
used to scale-in a cluster.
%
Since this protocol is fast and has low impact, it can also be used to
correctly partition records across servers.
%
In a distributed setting, partitioning becomes critical to
performance; it is well known that pre-partitioning records between
servers results in load imbalances, which significantly hurts
throughput~\cite{dynamo,slicer}.
%
Migration allows Shadowfax to
dynamically partition its hash space into arbitrary, fine-grained splits
and
avoid pre-partitioning.
%
Using load information available at runtime, it can first determine the ideal
way to split its hash space across servers.
%
It can then quickly migrate these splits
between them.
%
View validation helps too; a server can own many fine-grained splits and still
serve 100~Mops/s during normal operation.
